Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-06 ms.tex
FIXMEs

\subsection{White Light Curve Fit}
To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.  The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend. The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.  We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S_\mathrm{white}$ is a systematics model and $T_\mathrm{white}$ is a transit model. We used the same systematics model as \cite{kreidberg15b}, Equation 1.  To model the transit, we used the \texttt{batman} package \citep{kreidberg15a}.  The transit model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.

@@

\subsubsection{Star Spot Crossing}
In our initial analysis, we noticed a bump in the light curve during orbit three that could not be fit with our model. We attribute this feature to a star spot crossing event, as WASP-107 is an active star and spot crossings have been observed before \citep{dai17, mocnik17}. In our subsequent analysis, we gave the data in orbit three no weight in the fit. The amplitude of the spot crossing feature is 300 ppm.

@@

For the best fit white light curve, the root-mean-square (rms) residuals were 93 ppm (excluding the star spot crossing), which is somewhat larger than the expected shot noise of 50 ppm. We attribute the excess noise to loss of flux off the edge of the detector, which can occur if there is variation in the position or length of the spatial scan. There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi_2$ value of unity.  We then used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties \citep{foremanmackey13}.  The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.

\subsection{Spectroscopic Light Curve Fits}
We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.65 $\mu$m, shown in Figure\,\ref{fig:lc}. We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson??, kreidberg14a}. For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.  An advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction. There is a small chromatic effect due to the changing amplitude of the spot crossing feature with wavelength, but this effect is not detectable in our data.  As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model. The fit has just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.

The best fit light curves have a median reduced $\chi_2$ value of 1.16.  To ensure that we do not underestimate the uncertainty on the transit depths, we rescale the photometric uncertainties for all spectroscopic channels such that the $\chi_2$ values are unity. We performed an MCMC fit to the light curves with \texttt{emcee}. The median transit depths and $1\,\sigma$ uncertainties are given in Table FIXME.

We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum. We tested fitting the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (FIXME sigma), except offset by a constant value of FIXME ppm. We attribute the offset to the uncorrected star-spot crossing feature. Regardless of the cause, the offset does not affect our final analysis, because the planet's 1 bar radius is a free parameter in the atmospheric retrieval (see \S FIXME). We also explored fitting for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model predictions, and found that the `fitted limb darkening coefficients are consistent with the model.  We opted to fix the coefficients in our final analysis because the uncertainties on the transit depths are slightly smaller (about 10\%). We also checked whether the uncertainty in the stellar parameters significantly affects the PHOENIX model predictions, and found that changing the stellar

\subsubsection{Transit Depths Redder than FIXME $\mu$m}
The red edge of the transmission spectrum is of interest for our analysis because methane becomes the dominant absorber over water at wavelengths $> 1.6\,\mu$m. Unfortunately, we find that the spectroscopic light curves between FIXME $\mu$m are too poor in quality to accurately measure the transit depth. The residuals show evidence for correlated noise and the reduced $\chi_2$ values increase to FIXME.

@@

\begin{deluxetable}{c c c}
\tablecaption{WASP-107b transmission spectrum \label{transit_depths}}
\tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty}}
\startdata
1.133 & 0.020641 & 5.9e-05 \\
1.158 & 0.020733 & 5.5e-05 \\
1.184 & 0.020505 & 5.6e-05 \\
1.209 & 0.020498 & 5.4e-05 \\
1.235 & 0.020455 & 5.9e-05 \\
1.260 & 0.020492 & 5.0e-05 \\
1.285 & 0.020620 & 6.2e-05 \\
1.311 & 0.020739 & 5.0e-05 \\
1.336 & 0.020660 & 5.7e-05 \\
1.362 & 0.020858 & 4.8e-05 \\
1.387 & 0.020794 & 4.8e-05 \\
1.413 & 0.020888 & 5.2e-05 \\
1.438 & 0.020821 & 6.2e-05 \\
1.464 & 0.020691 & 5.1e-05 \\
1.489 & 0.020682 & 6.9e-05 \\
1.515 & 0.020679 & 6.7e-05 \\
1.540 & 0.020509 & 6.0e-05 \\
1.565 & 0.020480 & 6.4e-05 \\
1.591 & 0.020500 & 5.6e-05 \\
1.616 & 0.020514 & 6.5e-05 \\
\enddata
\end{deluxetable}

\section{Atmospheric Retrieval}
We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15b} to determine the basic atmospheric properties.   Briefly, the transmission spectrum solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.  We parametrize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.    We have subsequently upgraded the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database ($dv/v~10^6$) described in \cite{freedman14}. The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

@@

\section{Future Plans}
Combine data from Spitzer/HST
JWST GTO

>>>>

\subsection{White Light Curve}
To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.  The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend. The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.  We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S_\mathrm{white}$ is a systematics model and $T_\mathrm{white}$ is a transit model. We used the same systematics model as \cite{kreidberg15b}, Equation 1.  To model the transit, we used the \texttt{batman} package \citep{kreidberg15a}.  The transit model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.

@@

\subsubsection{Star Spot Crossing}
In our initial analysis, we noticed a bump in the light curve during orbit three that was not fit well with our model. We attribute this feature to a star spot crossing event, as WASP-107 is an active star and spot crossings have been observed before \citep{dai17, mocnik17}. In our subsequent analysis, we gave the data in orbit three no weight in the fit. The amplitude of the spot crossing feature is 300 ppm, as illustrated in Figure\,\ref{fig:lc}.

@@

For the best fit white light curve, the root-mean-square (rms) residuals were 93 ppm (excluding the star spot crossing), which is somewhat larger than the expected shot noise of 50 ppm. We attribute the excess noise to loss of flux off the edge of the detector, which can occur if there is variation in the position or length of the spatial scan. There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi^2$ value of unity.  We then used the Markov chain Monte Carlo (MCMC) algorithm to estimate parameter uncertainties \citep{foremanmackey13}.  The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$. FIXME: converged?

\subsection{Spectroscopic Light Curve Fits}
We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.65 $\mu$m, shown in Figure\,\ref{fig:lc}. We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson14c, kreidberg14a}. For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.  An advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction. There is a small chromatic effect due to the changing amplitude of the spot crossing feature with wavelength, but this effect is not detectable in our data.  As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model. The fit had just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.

The best fit light curves have a median reduced $\chi^2$ value of 1.16.  To ensure that we did not underestimate the uncertainty on the transit depths, we rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2$ values are unity. We performed an MCMC fit to the light curves with \texttt{emcee}.  The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:transit_depths}. FIXME: converged?

We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum. For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a constant offset of 350 ppm. We attribute the offset to the uncorrected star-spot crossing feature. Regardless of the cause, the offset does not affect our final analysis, because the planet's one-bar radius is a free parameter in the atmospheric retrieval (see \S\,\ref{sec:retrieval}). We also tried fitting for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model predictions, and found that the fitted limb darkening coefficients are consistent with the model.  We opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%. We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.

\subsubsection{Transit Depths Redder than 1.62 $\mu$m}
The red edge of the transmission spectrum is of interest for our analysis because methane is expected to be the dominant absorber over water at wavelengths greater than $1.6\,\mu$m. Unfortunately, we find that the our reddest spectroscopic light curves ($>1.62\mu$m) are too poor in quality to robustly measure the transit depth. The residuals exhibit correlated noise and the fits have reduced $\chi^2_\nu>2$. The transit depths are also sensitive to which method we use to fit for instrument systematics.  We therefore opt not to report transit depths redder than the 1.62 $\mu$m channel.

@@

\begin{deluxetable}{c c c}
\tablecaption{WASP-107b transmission spectrum \label{tab:transit_depths}}
\tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty}}
\startdata
1.133 & 0.020641 & 5.9e-05 \\
1.158 & 0.020733 & 5.5e-05 \\
1.184 & 0.020505 & 5.6e-05 \\
1.209 & 0.020498 & 5.4e-05 \\
1.235 & 0.020455 & 5.9e-05 \\
1.260 & 0.020492 & 5.0e-05 \\
1.285 & 0.020620 & 6.2e-05 \\
1.311 & 0.020739 & 5.0e-05 \\
1.336 & 0.020660 & 5.7e-05 \\
1.362 & 0.020858 & 4.8e-05 \\
1.387 & 0.020794 & 4.8e-05 \\
1.413 & 0.020888 & 5.2e-05 \\
1.438 & 0.020821 & 6.2e-05 \\
1.464 & 0.020691 & 5.1e-05 \\
1.489 & 0.020682 & 6.9e-05 \\
1.515 & 0.020679 & 6.7e-05 \\
1.540 & 0.020509 & 6.0e-05 \\
1.565 & 0.020480 & 6.4e-05 \\
1.591 & 0.020500 & 5.6e-05 \\
1.616 & 0.020514 & 6.5e-05 \\
\enddata
\end{deluxetable}

\section{Atmospheric Retrieval}
\label{sec:retrieval}
We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15b} to determine the basic atmospheric properties.   Briefly, the transmission spectrum solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.  We parametrize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.    We have subsequently upgraded the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database ($dv/v~10^6$) described in \cite{freedman14}. The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

@@

\section{Future Plans}
WASP-107b is slated to be observed with numerous facilities. Transits were recently observed with the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels. Spitzer eclipses are planned (Program FIXME; PI L. Kreidberg). In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations.

Combine data from Spitzer/HST
JWST GTO
